Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-01-05 manuscript/manuscript.Rmd
updated ms with climate sims

In contrast to individual-level data, population-level data arising from large-scale census efforts

Recently, @Freckleton2011, building on work by @Taylor2004, have proposed density-structured population models that focus on the transition of populations among discrete states, rather than the traditional approach of modeling the transitions of individuals. Such an approach could be extremely valuable because the data needed to parameterize density-structured population models is much easier, and less costly, to collect.

For example, using a density-structured approach, one could build a population model using a time series of annual plot-based censuses of species percent cover. However, a major assumption of the density-structured approach is that the aggregate dynamics of the population observed at coarse spatial resolution faithfully represent, and correspond to, the fates of individual plants. In other words, using a density-structured approach requires a leap of faith that important covariates (e.g., climate variables) at the level of the individual are captured adequately at the population level. If we seek to forecast the impacts of climate change on plant populations, then clearly this assumption requires testing.

Materials and Methods
---------------------
### Study site and data
Our demographic data comes from the Fort Keogh Livestock and Range Research Laboratory in eastern Montana's northern mixed prairie near Miles City, Montana, USA (46 deg. 19' N, 105 deg 48' W). The dataset is freely available on Ecological Archives [@Anderson2011], and interested readers should refer to the metadata therin for a complete description. The site is about 800 m above sea level and mean annual precipitation (1878-2009) is 334 mm, with most annual precipitaion falling from April through September. The site is grass dominated and, for the purposes of our study, we focus on the four most abundant graminoid species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE).

@@

```{r figure_4, dependson="plot-options", fig.cap="Simulated changes species' median cover caused by a 1% increase in observed precipitation or temperature. We used the median of simulated results because the average was highly influence by some extreme outliers.", fig.width=4.5, fig.height=4, cache=FALSE, results='hide'}
climBase <- readRDS("../montana/quadBM/simulations/climateChangeBase.rds")
climPpt <- readRDS("../montana/quadBM/simulations/climateChangePrecip.rds")
climTemp <- readRDS("../montana/quadBM/simulations/climateChangeTemperature.rds")
climBase$sim <- rep("Observed", nrow(climBase)) 
climPpt$sim <- rep("Precipitation", nrow(climBase)) 
climTemp$sim <- rep("Temperature", nrow(climBase)) 
climSimsQBM <- rbind(climBase, climPpt, climTemp)
# ggplot(climSimsQBM)+
#   geom_boxplot(aes(x=species, y=cover*100, fill=sim))+
#   scale_y_continuous(limits=c(0,10))
climSimsQBMavg <- ddply(climSimsQBM, .(sim, species), summarise,
                        avgCover = median(cover))
climSimQBMcast <- dcast(climSimsQBMavg, formula = species~sim)
climSimQBMcast$diffP <- with(climSimQBMcast, (Precipitation*100)-(Observed*100)) 
climSimQBMcast$diffT <- with(climSimQBMcast, (Temperature*100)-(Observed*100))
climSimQBMplot <- melt(climSimQBMcast)
ggplot(data=subset(climSimQBMplot, variable=="diffP"|variable=="diffT"))+
  geom_bar(aes(x=species, y=value, fill=variable), stat="identity", position="dodge")+
  scale_fill_manual(name="", labels=c("1% Precipitation Increase", "1% Temperature Increase"), values=c("grey25", "grey50"))+
  theme_bw()+
  theme(legend.position=c(0.6, 0.8))+
  ylab("Cover Change (%)")+
  xlab("Species")
```

>>>>

Aggregate measures of individual plant performance, such as those typically collected as part of large-scale census efforts, offer an alternative to detailed demographic data for modeling populations. Such population-level data will never match the precision of individual-level data,

Materials and Methods
---------------------
### Study site and data
Our demographic data comes from the Fort Keogh Livestock and Range Research Laboratory in eastern Montana's northern mixed prairie near Miles City, Montana, USA ($46^{\circ}$ 19' N, $105^{\circ}$ 48' W). The dataset is freely available on Ecological Archives [@Anderson2011], and interested readers should refer to the metadata therin for a complete description. The site is about 800 m above sea level and mean annual precipitation (1878-2009) is 334 mm, with most annual precipitaion falling from April through September. The site is grass dominated and, for the purposes of our study, we focus on the four most abundant graminoid species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE).

@@

```{r figure_4, dependson="plot-options", fig.cap="Simulated changes species' median cover caused by a 1% increase in observed precipitation or temperature. We used the median of simulated results because the average was highly influence by some extreme outliers.", fig.width=4.5, fig.height=4, cache=FALSE, results='hide'}
climBase <- readRDS("../montana/quadBM/simulations/climateChangeBase.rds")
climPpt <- readRDS("../montana/quadBM/simulations/climateChangePrecip.rds")
climTemp <- readRDS("../montana/quadBM/simulations/climateChangeTemperature.rds")
climBoth <- readRDS("../montana/quadBM/simulations/climateChangePptTmp.rds")
climBase$sim <- rep("Observed", nrow(climBase)) 
climPpt$sim <- rep("Precipitation", nrow(climBase)) 
climTemp$sim <- rep("Temperature", nrow(climBase)) 
climBoth$sim <- rep("Both", nrow(climBase)) 
climSimsQBM <- rbind(climBase, climPpt, climTemp, climBoth)
# ggplot(climSimsQBM)+
#   geom_boxplot(aes(x=species, y=cover*100, fill=sim))+
#   scale_y_continuous(limits=c(0,10))
climSimsQBMavg <- ddply(climSimsQBM, .(sim, species), summarise,
                        avgCover = median(cover))
climSimQBMcast <- dcast(climSimsQBMavg, formula = species~sim)
climSimQBMcast$diffP <- with(climSimQBMcast, (Precipitation*100)-(Observed*100)) 
climSimQBMcast$diffT <- with(climSimQBMcast, (Temperature*100)-(Observed*100))
climSimQBMcast$diffB <- with(climSimQBMcast, (Both*100)-(Observed*100))
climSimQBMplot <- melt(climSimQBMcast)
ggplot(data=subset(climSimQBMplot, variable=="diffP"|variable=="diffT"|variable=="diffB"))+
  geom_bar(aes(x=species, y=value, fill=variable), stat="identity", position="dodge", color="white")+
  scale_fill_manual(name="", labels=c("1% Precipitation Increase", "1% Temperature Increase", "1% Temp. & Ppt. Increase"), values=c("grey25", "grey50", "grey75"))+
  theme_bw()+
  theme(legend.position=c(0.6, 0.8))+
  ylab("Cover Change (%)")+
  xlab("Species")
```

### 2015-01-05 manuscript/manuscript.tex
updated ms with climate sims

In contrast to individual-level data, population-level data arising from
large-scale census efforts

Recently, Freckleton et al. (2011), building on work by Taylor and
Hastings (2004), have proposed density-structured population models that
focus on the transition of populations among discrete states, rather
than the traditional approach of modeling the transitions of
individuals. Such an approach could be extremely valuable because the
data needed to parameterize density-structured population models is much
easier, and less costly, to collect.

For example, using a density-structured approach, one could build a
population model using a time series of annual plot-based censuses of
species percent cover. However, a major assumption of the
density-structured approach is that the aggregate dynamics of the
population observed at coarse spatial resolution faithfully represent,
and correspond to, the fates of individual plants. In other words, using
a density-structured approach requires a leap of faith that important
covariates (e.g., climate variables) at the level of the individual are
captured adequately at the population level. If we seek to forecast the
impacts of climate change on plant populations, then clearly this
assumption requires testing.

@@

Our demographic data comes from the Fort Keogh Livestock and Range
Research Laboratory in eastern Montana's northern mixed prairie near
Miles City, Montana, USA (46 deg. 19' N, 105 deg 48' W). The dataset is
freely available on Ecological Archives (Anderson et al. 2011), and
interested readers should refer to the metadata therin for a complete
description. The site is about 800 m above sea level and mean annual
precipitation (1878-2009) is 334 mm, with most annual precipitaion
falling from April through September. The site is grass dominated and,
for the purposes of our study, we focus on the four most abundant
graminoid species: \emph{Bouteloua gracilis} (BOGR),
\emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii} (PASM), and
\emph{Poa secunda} (POSE).

@@

@@

Taylor, C. M., and A. Hastings. 2004. Finding optimal control strategies
for invasive species: a density-structured model for Spartina
alterniflora. Journal of Applied Ecology 41:1049--1057.

>>>>

Aggregate measures of individual plant performance, such as those
typically collected as part of large-scale census efforts, offer an
alternative to detailed demographic data for modeling populations. Such
population-level data will never match the precision of individual-level
data,

@@

Our demographic data comes from the Fort Keogh Livestock and Range
Research Laboratory in eastern Montana's northern mixed prairie near
Miles City, Montana, USA ($46^{\circ}$ 19' N, $105^{\circ}$ 48' W). The
dataset is freely available on Ecological Archives (Anderson et al.
2011), and interested readers should refer to the metadata therin for a
complete description. The site is about 800 m above sea level and mean
annual precipitation (1878-2009) is 334 mm, with most annual
precipitaion falling from April through September. The site is grass
dominated and, for the purposes of our study, we focus on the four most
abundant graminoid species: \emph{Bouteloua gracilis} (BOGR),
\emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii} (PASM), and
\emph{Poa secunda} (POSE).

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Climate detections
\item
  Limitations:
\end{enumerate}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Sample size at quadrat level
\item
  are these species inherently sensitive to climate (compare to Chu and
  Adler in review -- say why we chose Montana site -- most sensitive to
  climate variables)
\item
  we know recruitment is important, and this is something that may be
  easily missed at quad level
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Implications
\end{enumerate}

@@
